---
title: "Collaborator Analysis"
format: html
author: "Andie Norton"
---

## Exercise: Explore, Clean, and Wrangle Data

#### Load necessary documents
```{r}
library(readr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(here)
library(janitor)
```

#### bring in the data!
```{r}
lobster_traps_raw <- read_csv("../data/Lobster_Trap_Counts_All_Years_20210519.csv")
# NOTE: Fishing season is a char col, will need to separate into szn_start and szn_end by "-"
## also segment start and segment end are both chr because they are in DDM:  divide the minutes by 
## 60 and add the result to the degrees

## NOTE2: Authors used "-99999" to denote an NA value, this will need to be changed

```

#### Clean the data
```{r}

lobster_traps <- lobster_traps_raw %>% clean_names() %>% #clean names up
  mutate(traps = na_if(traps, -99999)) %>% select(-notes) # fix NA values, remove notes col
head(lobster_traps)

```

#### Remove Naples Reef (NAPL)
```{r}
traps_noNAPL <- lobster_traps %>% filter(site != "NAPL")
unique(lobster_traps_noNAPL$site)
```

#### Filter for Carpinteria Reef and traps > 20
```{r}
traps_CARP20 <- lobster_traps %>% filter(site == "CARP" & traps > 20)
```

#### Find the maximum number of commercial trap floats group by SITE and MONTH.
```{r}
traps_max <- lobster_traps %>% group_by(site, month) %>% filter(traps == max(traps, na.rm=TRUE))
```

## Exercise: Data Visualization

#### traps faceted by Year
```{r}
ggplot(data = lobster_traps, aes(x = traps)) +
    geom_histogram() +
    facet_wrap( ~ year)
    
```

#### line graph of the number of total lobster commercial traps observed (y-axis) by year (x-axis) 
```{r}
lobsters_traps_summarize
```

