---
title: "Lobster Report"
format: html
---

## Summary: Data on abundance, size and fishing pressure of California spiny lobster are collected along the mainland coast of the Santa Barbara Channel. Sampling began in 2012 and is ongoing {citation}.


## Owner analysis and visualizations
## Exercise: Explore, Clean, and Wrangle Data

### Loading libraries

```{r}
#| warning: false

library(readr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(here)
```

### Reading the data

```{r}
lobster_abundance <- read_csv(here("data/Lobster_Abundance_All_Years_20220829.csv"))
```

```{r}
head(lobster_abundance)
```
```{r}
colnames(lobster_abundance)
```


### Convert missing values using mutate() and na_if()
```{r}
lobster_abundance <- lobster_abundance %>% 
    mutate(SIZE_MM = na_if(SIZE_MM, -99999))
```

### filter() practice
```{r}
not_napl <- lobster_abundance %>% 
  filter(SITE != "NAPL")

aque_70mm <- lobster_abundance %>% 
  filter(SITE == "AQUE" & SIZE_MM >= 70)

max_lobster <- lobster_abundance %>% 
  group_by(SITE, MONTH) %>% 
  summarize(MAX_LENGTH = max(SIZE_MM, na.rm = TRUE))
```

### Visualization
```{r}
p_hist_site <- lobster_abundance %>%
  drop_na(SIZE_MM) %>%
  ggplot(aes(x = SIZE_MM)) +
  geom_histogram(bins = 30, alpha = 0.8, color = "white") +
  facet_wrap(~ SITE, ncol = 3) +
  labs(
    title = "California spiny lobster carapace length by site",
    subtitle = "Mainland Santa Barbara Channel LTER lobster monitoring",
    x = "Carapace length (mm)",
    y = "Number of lobsters"
  ) +
  scale_x_continuous(breaks = seq(0, 200, by = 20)) +
  theme_light() +
  theme(
    plot.title = element_text(face = "bold"),
    strip.text = element_text(face = "bold"),
    panel.grid.minor = element_blank(),
    axis.title = element_text(face = "bold")
  )

p_hist_site
```


