---
title: "Owner Analysis"
author: "Valentina Ekimova"
format: html
---

## Exercise: Explore, Clean, and Wrangle Data

### Loading libraries

```{r}
#| warning: false

library(readr)
library(dplyr)
library(ggplot2)
library(tidyr)
library(here)
```

### Reading the data

```{r}
lobster_abundance <- read_csv(here("data/Lobster_Abundance_All_Years_20220829.csv"))
```

```{r}
head(lobster_abundance)
```
```{r}
colnames(lobster_abundance)
```


### Convert missing values using mutate() and na_if()
```{r}
lobster_abundance <- lobster_abundance %>% 
    mutate(SIZE_MM = na_if(SIZE_MM, -99999))
```

### filter() practice
```{r}
not_napl <- lobster_abundance %>% 
  filter(SITE != "NAPL")

aque_70mm <- lobster_abundance %>% 
  filter(SITE == "AQUE" & SIZE_MM >= 70)

max_lobster <- lobster_abundance %>% 
  group_by(SITE, MONTH) %>% 
  summarize(MAX_LENGTH = max(SIZE_MM, na.rm = TRUE))
```

## Exercise: Data Visualization

### To upload libraries for visualization
```{r}
library(readr)
library(dplyr)
library(tidyr)
library(ggplot2)
```

### Multi-panel plot of lobster carapace length
```{r}
ggplot(data = lobster_abundance, aes(x = SIZE_MM)) +
  geom_histogram(bins = 30) +
  facet_wrap(~SITE)
```
### Graph of the number of total lobsters observed
```{r}
# Create a new dataset subset called lobsters_summarize

lobsters_summarize <- lobster_abundance %>% 
  group_by(SITE, YEAR) %>% 
  summarize(COUNT = n())

# Create a line graph
ggplot(data = lobsters_summarize, aes(x = YEAR, y = COUNT)) +
  geom_line(aes(color = SITE)) 

ggplot(data = lobsters_summarize, aes(x = YEAR, y = COUNT)) +
  geom_point(aes(color = SITE)) +
  geom_line(aes(color = SITE)) 
```
### Create a bar graph that shows the amount of small and large sized carapace lobsters at each SITE
```{r}
# Create a new dataset subset called lobsters_size_lrg

lobster_size_lrg <- lobster_abundance %>%
    filter(YEAR %in% c(2019, 2020, 2021)) %>%
    mutate(SIZE_BIN = if_else(SIZE_MM <= 70, true = "small", false = "large")) %>%
    group_by(SITE, SIZE_BIN) %>%
    summarize(COUNT = n()) %>%
    drop_na()

# bar plot
ggplot(data = lobster_size_lrg, aes(x = SITE, y = COUNT, fill = SIZE_BIN)) +
    geom_col()

# dodged bar plot
ggplot(data = lobster_size_lrg, aes(x = SITE, y = COUNT, fill = SIZE_BIN)) +
    geom_col(position = "dodge")

```
### Creating a beautiful graph
```{r}
p_hist_site <- lobster_abundance %>%
  drop_na(SIZE_MM) %>%
  ggplot(aes(x = SIZE_MM)) +
  geom_histogram(bins = 30, alpha = 0.8, color = "white") +
  facet_wrap(~ SITE, ncol = 3) +
  labs(
    title = "California spiny lobster carapace length by site",
    subtitle = "Mainland Santa Barbara Channel LTER lobster monitoring",
    x = "Carapace length (mm)",
    y = "Number of lobsters"
  ) +
  scale_x_continuous(breaks = seq(0, 200, by = 20)) +
  theme_light() +
  theme(
    plot.title = element_text(face = "bold"),
    strip.text = element_text(face = "bold"),
    panel.grid.minor = element_blank(),
    axis.title = element_text(face = "bold")
  )

p_hist_site
```
### Saving graph
```{r}

ggsave(
  filename = here("figs", "hist_carapace_length_by_site.png"),
  plot = p_hist_site,
  width = 10,
  height = 7,
  dpi = 300
)

```

